---
title: "Assignment 3"
author: "Jacob Fabian"
date: "2023-04-27"
output:
  pdf_document: default
  html_document: default
---

#Part A

#Question 1
-What is the difference between SVM with hard margin and soft margin?
A hard margin is a SVM model with very strict margins that will not allow the model to violate them. A Soft margin is the opposite, a very lenient model with margins that can be violated to a cetain extent if it means a better split.

#Question 2 
-What is the role of the cost parameter, C, in SVM (with soft margin) classifiers?

The cost parameter tells the model how strict or lenient the model can be with the constraints. The larger C means the stricter the model is with the constraints. A smaller C allows for the model to be more lenient with constraints and even ignoring the constraints when necessary, this will lead to a large margin. A larger C will be more strict and not all the model to break the constraints, leading to a narrow margin. If C = infinity, then the constraints are a hard margin. Soft margins provide a trade off between the number of violations allowed and the margins.

#Question 3
-Will the following perceptron be activated (2.8 is the activation threshold)?

11.1*-0.2 = -2.22 then you add the 0.08. -2.22 + 0.08 = -2.14. The perceptron will not be activated since the result is -2.14.

#Question 4
-What is the role of alpha, the learning rate in the delta rule?

In the role of alpha, the learning rate sets how to change the weights (w) as a function of the inputs (x), outputs (y) and the target (t). A higher alpha or learning rate, will make the changes faster to the weight. A lower alpha will result in quicker change to the weight. By using a higher learning rate in the beginning, we will get closer to the target. Switching to a smaller learning rate will then smoothly reach the optimal value. 


#Part B

Loading Libraries

```{r}
library(ISLR)
library(dplyr)
library(glmnet)
library(caret)
library(kernlab)
```

Build a linear SVM regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). Hint: use caret train() with method set to “svmLinear”. What is the R-squared of the model?

```{r carseats}
Carseats_Filtered <- Carseats %>% select("Sales", "Price", "Advertising","Population","Age","Income","Education") 
```

#Question 1
- Build a linear SVM regression model to predict Sales based on all other attributes ("Price", "Advertising", "Population", "Age", "Income" and "Education"). Hint: use caret train() with method set to  “svmLinear”. What is the R-squared of the model?

```{r}
trctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 3)
set.seed(23)
svm_Linear <- train(Sales~., data = Carseats_Filtered, method = "svmLinear",
                    trControl=trctrl,
                    preProcess = c("center", "scale"),
                    tuneLength = 10)
svm_Linear
```
The R^2 of the model is = 0.369034 


#Question 2
-Customize the search grid by checking the model’s performance for C parameter of 0.1,.5,1 and 10 using 2 repeats of 5-fold cross validation. 

```{r}
grid <- expand.grid(C = c(0.1,0.5,1,10))
trctrl2 <- trainControl(method = "repeatedcv", number = 5, repeats = 2)
svm_Linear_Grid <- train(Sales~., data = Carseats_Filtered, method = "svmLinear",
                         trControl=trctrl2,
                         preProcess = c("center", "scale"),
                         tuneGrid = grid,
                         tuneLength = 10)
svm_Linear_Grid
```
The best C is 0.5

#Question 3
-Train a neural network model to predict Sales based on all other attributes ("Price",
"Advertising", "Population", "Age", "Income" and "Education"). Hint: use caret train() withmethod set to “nnet”. What is the R-square of the model with the best hyper parameters (using default caret search grid) – hint: don’t forget to scale the data.

```{r}
set.seed(23)
numfolds <- trainControl(method = 'LOOCV', verboseIter = FALSE)
nnet_Cars <- train(Sales~., data = Carseats_Filtered, method = "nnet",
                    preProcess = c("center", "scale"),
                   trControl = numfolds)
nnet_Cars
```
The most optimal model using RMSE has a size = 1 and a decay = 1e-04.
The Rsquared for this model = "NAN". 
The closest Rsquared is = 0.32235623 for a model with a decay of 1e-01 and a RMSE of 7.082396.

#Question 4
-Consider the following input: Sales=9, Price=6.54, Population=124, Advertising=0, Age=76, Income= 110, Education=10. What will be the estimated Sales for this record using the above neuralnet model?

```{r}
Sales <- c(9)
Price <- c(6.54)
Population <- c(124)
Advertising <- c(0)
Age <- c(76)
Income <- c(110)
Education <- c(10)
Test <- data.frame(Sales, Price, Population, Advertising, Age, Income, Education)
```

#Time to predict using our network.

```{r}
Pred_sales <- predict(nnet_Cars, Test)
Pred_sales
```
After building the Neural net model, the model predicts that "1" sale will take place. In the previous assignment, "9.5" sales would occur at this record. Using the Keras package may provide a better outcome. Building a model with "keras_model_sequential" may create a more adjustable neural network model. 

